Supplementary Information: Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments

### 2025-04-05 doc/manuscript_hkk_drought_sensitivity.Rmd
condensing abstract

# Abstract (150 words) {-}
The frequency and intensity of droughts are increasing across forests worldwide, altering tree demography and its contribution to the land carbon sink. 
In seasonally dry tropical forests, species adaptations to strong dry seasons might confer drought resistance, but this resistance may be variable with individual microenvironments and across drought events. 
We analysed a unique data record - manual dendrometer band measurements over 15 years on 1820 trees across 30 species in a seasonally dry tropical forest in <!---Huai Kha Khaeng, ---->Thailand, <!--using Bayesian regressions in a causal modelling framework--->to assess the drought sensitivity of tree growth in two climatically distinct droughts. 
Species and individuals exhibited a wide range of drought sensitivities within a drought event, and inconsistent responses across these two droughts. 
Deciduous species were more sensitive than evergreen species in the wet-season drought, and this difference was amplified in wetter microclimates. 
Although exposed trees were more sensitive, access to water through larger size or topographic location buffered growth declines. 
Heterogeneous responses of species and individuals to droughts indicate potential spatiotemporal insurance effects in diverse forests in the face of increased climate anomalies.

>>>>

# Abstract (150 words) {-}
Droughts are increasingly impacting forests worldwide, altering tree demography and its contribution to the land carbon sink. 
In tropical dry forests, species adaptations to strong dry seasons (e.g., deciduousness) might confer drought resistance, but this resistance may be variable across microenvironments and drought events. 
We analysed 15 years of dendrometer band measurements on 1820 trees across 30 species in a seasonally dry tropical forest in <!---Huai Kha Khaeng, ---->Thailand <!--using Bayesian regressions in a causal modelling framework--->to assess the drought sensitivity of tree growth in two climatically distinct droughts. 
Species and individuals exhibited a wide range of drought sensitivities within a drought event, and inconsistent responses across these two droughts. 
Deciduous species were more sensitive than evergreen species in the wet-season drought, and this difference was amplified in wetter microclimates. 
Although exposed trees were more sensitive, access to water through larger size or topographic location buffered growth declines. 
Heterogeneous responses of species and individuals to droughts indicate potential spatiotemporal insurance effects in diverse forests in the face of increased climate anomalies.
